---
title: "Initial Korean covid 19"
author: "Edward Gomez"
date: "fall 2025"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?










## Graphical results - use ggplot commands to do this  

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This data seems to show that Covid is killing men at a nearly 2:1 ratio when compared to women.



  





  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

This data shows that 56% of those who survived the disease were women, while 44% were men. The data also shows that the percentage of women who died was 1.4% and the percentage of men who died was nearly double, at 3%.
## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The data shows a p-value of 0.0009, a value small enough to show that there a statistical difference between the two groups. The data also shows an odds ratio of 2.3, which means there is a 2.3 times greater chance that a male with covid will pass away when compared to females. 

## Conclusions

 I decided to use the Cat_Cat file to help me solve solve this problem since it is a categorical~categorical problem. Plotting the data I found that there was a higher proportion of women participating and a higher proportion of men dying in the study. The data shows that 1.4 percent of women died compared to 3 percent of men dying. The data yielded a p-value of 0.0009 and an odds ratio of 2.3, which shows that men died at a rate 2.3 greater than that of women. 
